Covariate analysis of MAYO reprocessed counts (with TMM normalisation)

### 2017-01-30 code/Rmd/MSSM_RNASeq_Reprocessed_TMM.Rmd
Minor bug and sanity fixes

Clustering of initial normalised data (with NULL design)
```{r decompse.normalise.data1, fig.height=14, fig.width=7, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(VOOM.GENE_EXPRESSION$E, scale. = T, center = T)

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=factor(CDR), color=as.factor(BrodmannArea), size=as.numeric(RIN)))
p <- p + theme_bw() + theme(legend.position="top") +scale_shape_manual(values = 1:7)
# p <- p + geom_text(aes(label= sampleIdentifier), size=4, hjust=0)
pl[[1]] <- p

p <- ggplot(plotdata, aes(x=PC2, y=PC3))
p <- p + geom_point(aes(shape=factor(CDR), color=as.factor(BrodmannArea), size=as.numeric(RIN)))
p <- p + theme_bw() + theme(legend.position="top") +scale_shape_manual(values = 1:7)
# p <- p + geom_text(aes(label= sampleIdentifier), size=4, hjust=0)
pl[[2]] <- p

@@

tmp <- paste('Using following covariates in the final model:', modelStr, 'and Donor_ID as random effects')
```
`r tmp`
### Sanity Check
Residualise significant covariates using a linear model and find correlation between PCA of residuals with covariates
```{r residualSigCovars.ALL, fig.width=25, fig.height=12}
residualSigCovars[["PC_res"]][[2]]$plotData
```
Adding "BrodmannArea.CDR" back to the residuals
```{r addVarsBack}
varsToAddBackIn <-  grep("BrainRegion.Diagnosis", colnames(FIT$coefficients), value = T)
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  t(data.matrix(DM1$design[,varsToAddBackIn]) %*% t(FIT$coefficients[,varsToAddBackIn]))
```

### Clustering residual gene expression
```{r decompse.normalise.data3, fig.height=10, fig.width=9, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION, scale.=T, center = T)

@@

pl = list()
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=factor(CDR), color=as.factor(BrodmannArea), size=as.numeric(RIN)))
p <- p + theme_bw() + theme(legend.position="top") +scale_shape_manual(values = 1:7)
# p <- p + geom_text(aes(label= sampleIdentifier), size=4, hjust=0)
pl[[1]] <- p

p <- ggplot(plotdata, aes(x=PC2, y=PC3))
p <- p + geom_point(aes(shape=factor(CDR), color=as.factor(BrodmannArea), size=as.numeric(RIN)))
p <- p + theme_bw() + theme(legend.position="top") +scale_shape_manual(values = 1:7)
# p <- p + geom_text(aes(label= sampleIdentifier), size=4, hjust=0)
pl[[2]] <- p

@@

# Get differnetial expression
DE = lapply(1:84, function(i, FIT.CONTR, VOOM.ADJUSTED.GENE_EXPRESSION){
  topTable(FIT.CONTR, coef=i, number = dim(VOOM.ADJUSTED.GENE_EXPRESSION$E)[1])
}, FIT.CONTR, VOOM.ADJUSTED.GENE_EXPRESSION)
names(DE) = gsub('BrodmannArea.CDR','',colnames(contrast))

@@

# Store differential expression results
write.table(DE, file = 'MSBBDiffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('MSBBDiffExpression.tsv', name = 'Differential Expression Analysis', parentId = CODE$properties$id)
DEXP_OBJ = synStore(DEXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                    executed = thisFile, activityDescription = activityDescription)

>>>>

Clustering of initial normalised data (with NULL design)
```{r decompse.normalise.data1, fig.height=10, fig.width=10, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(VOOM.GENE_EXPRESSION$E, scale. = T, center = T)

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=factor(CDR), color=as.factor(BrodmannArea), size=as.numeric(RIN)))
p <- p + theme_bw() + theme(legend.position="right") +scale_shape_manual(values = 1:7)
# p <- p + geom_text(aes(label= sampleIdentifier), size=4, hjust=0)
pl[[1]] <- p

p <- ggplot(plotdata, aes(x=PC2, y=PC3))
p <- p + geom_point(aes(shape=factor(CDR), color=as.factor(BrodmannArea), size=as.numeric(RIN)))
p <- p + theme_bw() + theme(legend.position="right") +scale_shape_manual(values = 1:7)
# p <- p + geom_text(aes(label= sampleIdentifier), size=4, hjust=0)
pl[[2]] <- p

@@

tmp <- paste('Using following covariates in the final model:', modelStr, 'and Donor_ID as random effects')
```
`r tmp`
### Sanity Check
Residualise significant covariates using a linear model and find correlation between PCA of residuals with covariates
```{r residualSigCovars.ALL, fig.width=25, fig.height=12}
residualSigCovars[["PC_res"]][[2]]$plotData
```
Adding "BrodmannArea.CDR" back to the residuals
```{r addVarsBack}
varsToAddBackIn <-  grep("BrodmannArea.CDR", colnames(FIT$coefficients), value = T)
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  t(data.matrix(DM1$design[,varsToAddBackIn]) %*% t(FIT$coefficients[,varsToAddBackIn]))
```

### Clustering residual gene expression
```{r decompse.normalise.data3, fig.height=10, fig.width=10, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION, scale.=T, center = T)

@@

pl = list()
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=factor(CDR), color=as.factor(BrodmannArea), size=as.numeric(RIN)))
p <- p + theme_bw() + theme(legend.position="right") +scale_shape_manual(values = 1:7)
# p <- p + geom_text(aes(label= sampleIdentifier), size=4, hjust=0)
pl[[1]] <- p

p <- ggplot(plotdata, aes(x=PC2, y=PC3))
p <- p + geom_point(aes(shape=factor(CDR), color=as.factor(BrodmannArea), size=as.numeric(RIN)))
p <- p + theme_bw() + theme(legend.position="right") +scale_shape_manual(values = 1:7)
# p <- p + geom_text(aes(label= sampleIdentifier), size=4, hjust=0)
pl[[2]] <- p

@@

# Get differnetial expression
DE = lapply(1:84, function(i, FIT.CONTR, VOOM.ADJUSTED.GENE_EXPRESSION){
  topTable(FIT.CONTR, coef=i, number = dim(VOOM.ADJUSTED.GENE_EXPRESSION$E)[1], confint = T)
}, FIT.CONTR, VOOM.ADJUSTED.GENE_EXPRESSION)
names(DE) = gsub('BrodmannArea.CDR','',colnames(contrast))

@@

# Store differential expression results
write.table(DE1, file = 'MSBBDiffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('MSBBDiffExpression.tsv', name = 'Differential Expression Analysis', parentId = CODE$properties$id)
DEXP_OBJ = synStore(DEXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                    executed = thisFile, activityDescription = activityDescription)

### 2017-01-30 code/Rmd/MSSM_RNASeq_Reprocessed_TMM_BinnedCDR.Rmd
Minor bug and sanity fixes

DE1 = plyr::ldply(DE, .id = 'Comparison') %>%
  dplyr::mutate(Comparison1 = Comparison) %>%
  tidyr::separate(Comparison1, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('BrodmannArea','From.Dx'), sep = 5) %>%
  tidyr::separate(to.state, c('BrodmannArea','To.Dx'), sep = 5) %>%
  tidyr::unite(Dx, From.Dx, To.Dx, sep = '-')

p <- ggplot(DE1, aes(x = logFC, y = -log10(adj.P.Val))) + geom_point()
p <- p + geom_hline(yintercept = -log10(0.05), color = 'red')
p <- p + geom_vline(xintercept = log2(1.2), color = 'red')
p <- p + geom_vline(xintercept = -log2(1.2), color = 'red')
p <- p + facet_grid(BrodmannArea~.+Dx)
p

tmp = DE1 %>% 
  dplyr::filter(adj.P.Val <= 0.05) %>%
  group_by(BrodmannArea, Dx) %>%
  summarise(count = length(unique(genes))) %>%
  tidyr::spread(BrodmannArea, count)
kable(tmp)
```

>>>>

DE1 = plyr::ldply(DE, .id = 'Comparison') %>%
  dplyr::mutate(Comparison1 = Comparison) %>%
  tidyr::separate(Comparison1, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('BrodmannArea','From.Dx'), sep = 5) %>%
  tidyr::separate(to.state, c('BrodmannArea','To.Dx'), sep = 5) %>%
  tidyr::unite(Dx, From.Dx, To.Dx, sep = '-') %>%
  dplyr::select(-Comparison) %>%
  dplyr::rename(Region = BrodmannArea, Comparison = Dx, gene.id = genes) %>%
  dplyr::mutate(Region = factor(Region, labels = c('FP', 'STG', 'PHG', 'IFG')),
                Study = 'MSBB') %>%
  left_join(backgroundGenes) %>%
  left_join(Ensemble2HGNC) %>%
  dplyr::select(-position)

p <- ggplot(DE1, aes(x = logFC, y = -log10(adj.P.Val))) + geom_point()
p <- p + geom_hline(yintercept = -log10(0.05), color = 'red')
p <- p + geom_vline(xintercept = log2(1.2), color = 'red')
p <- p + geom_vline(xintercept = -log2(1.2), color = 'red')
p <- p + facet_grid(Region~.+Comparison)
p

tmp = DE1 %>% 
  dplyr::filter(adj.P.Val <= 0.05) %>%
  group_by(Region, Comparison) %>%
  summarise(count = length(unique(genes))) %>%
  tidyr::spread(Region, Comparison)
kable(tmp)
```

### 2017-01-30 code/Rmd/MSSM_RNASeq_Reprocessed_TMM_SVA.Rmd
Minor bug and sanity fixes

### Clustering residual gene expression
```{r decompse.normalise.data3, fig.height=10, fig.width=9, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION, scale.=T, center = T)

>>>>

### Clustering residual gene expression
```{r decompse.normalise.data3, fig.height=10, fig.width=10, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION, scale.=T, center = T)

### 2017-01-30 code/Rmd/MSSM_RNASeq_Reprocessed_TMM_SVA_BinnedCDR.Rmd
Minor bug and sanity fixes

DE1 = plyr::ldply(DE, .id = 'Comparison') %>%
  dplyr::mutate(Comparison1 = Comparison) %>%
  tidyr::separate(Comparison1, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('BrodmannArea','From.Dx'), sep = 5) %>%
  tidyr::separate(to.state, c('BrodmannArea','To.Dx'), sep = 5) %>%
  tidyr::unite(Dx, From.Dx, To.Dx, sep = '-')

p <- ggplot(DE1, aes(x = logFC, y = -log10(adj.P.Val))) + geom_point()
p <- p + geom_hline(yintercept = -log10(0.05), color = 'red')
p <- p + geom_vline(xintercept = log2(1.2), color = 'red')
p <- p + geom_vline(xintercept = -log2(1.2), color = 'red')
p <- p + facet_grid(BrodmannArea~.+Dx)
p

tmp = DE1 %>% 
  dplyr::filter(adj.P.Val <= 0.05) %>%
  group_by(BrodmannArea, Dx) %>%
  summarise(count = length(unique(genes))) %>%
  tidyr::spread(BrodmannArea, count)
kable(tmp)
```

>>>>

DE1 = plyr::ldply(DE, .id = 'Comparison') %>%
  dplyr::mutate(Comparison1 = Comparison) %>%
  tidyr::separate(Comparison1, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('BrodmannArea','From.Dx'), sep = 5) %>%
  tidyr::separate(to.state, c('BrodmannArea','To.Dx'), sep = 5) %>%
  tidyr::unite(Dx, From.Dx, To.Dx, sep = '-') %>%
  dplyr::select(-Comparison) %>%
  dplyr::rename(Region = BrodmannArea, Comparison = Dx, gene.id = genes) %>%
  dplyr::mutate(Region = factor(Region, labels = c('FP', 'STG', 'PHG', 'IFG')),
                Study = 'MSBB') %>%
  left_join(backgroundGenes) %>%
  left_join(Ensemble2HGNC) %>%
  dplyr::select(-position)

p <- ggplot(DE1, aes(x = logFC, y = -log10(adj.P.Val))) + geom_point()
p <- p + geom_hline(yintercept = -log10(0.05), color = 'red')
p <- p + geom_vline(xintercept = log2(1.2), color = 'red')
p <- p + geom_vline(xintercept = -log2(1.2), color = 'red')
p <- p + facet_grid(Region~.+Comparison)
p

tmp = DE1 %>% 
  dplyr::filter(adj.P.Val <= 0.05) %>%
  group_by(Region, Comparison) %>%
  summarise(count = length(unique(gene.id))) %>%
  tidyr::spread(Region, count)
kable(tmp)
```

### 2017-01-30 code/Rmd/Mayo_RNASeq_Reprocessed_TMM.Rmd
Minor bug and sanity fixes

---
title: "Covariate analysis of MAYO reprocessed counts (with TMM and SVA normalisation)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knit2synapse)
library(synapseClient)

>>>>

---
title: "Covariate analysis of MAYO reprocessed counts (with TMM normalisation)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knit2synapse)
library(synapseClient)
